---
title: "Tax Havens and Secret Finance"
author: "Kiersten Hileman"
date: "04-07-23"
output:
  xaringan::moon_reader:
    css: [default, duke-blue, fc-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)


library(haven)
library(knitr)
library(tidyverse)
library(scales)
library(ggplot2)
library(patchwork)
library(dplyr)
library(ggplot2)
library(ggrepel)

df = read.csv('CTHI_2021_Main_Results.csv')
df1 = read.csv('fsi_2022_main_results.csv') 
colnames(df1)[2] = 'ISO.2'
colnames(df1)[13] = 'FSI Rank'
df2 <- full_join(df, df1,
          by = join_by("ISO.2"))
na.omit(df2)
df3 <- df2 %>% distinct()
colnames(df3)[21] = 'FSI Score'
```





## Data Sources  
### Tax Justice Network (TJN)
###Illicit Financial Flows Vulnerabilty Tracker
* TJN collects data from the World Bank and IMF on 207 countries but filters out a large amount of these based on global economic weight 
* Two main measures used here: Tax Haven Score and Financial Secrecy Ranking
* Tax Haven Scores is calculated using 4 different categories of law that indicate ability to evade taxes:   
    * Lowest available corporate income tax  
    * Loopholes and gaps, 
    * Transparency
    * Anti-avoidance  
    * Double tax treaty aggressiveness. 
    * 0 meaning does not allow for corporate tax abuse and 100 correlates with unrestricted corporate tax abuse. 
* Financial Secrecy Ranking is calculated across 151 jurisdictions based on 20 indicators of financial secrecy that revolve around governments' anti-money laundering laws and tax haven status
    * Weighted against Global Scale Weight based on economy size where 100 is perfectly unrestrained secrecy and 0 meaning no financial secrecy is allowed. 


---
class: inverse, center, middle
## Taxen Haven Rankings
```{r, echo=FALSE, warnings = FALSE, }
p <- ggplot(data = df, aes(x = Rank, y = Haven.Score, label = ISO.2)) +
  geom_point(color = case_when(
             df$Rank >= 65 ~ 'dodgerblue',
             df$Rank <= 5 ~ 'forestgreen',
             TRUE ~ 'slateblue1'), 
               size = 3, alpha = 0.8) +
  geom_text_repel(data          = subset(df, Rank >= 65),
                  nudge_y       = 32 - subset(df, Rank >= 65)$Rank,
                  size          = 4,
                  box.padding   = 1.5,
                  point.padding = 0.5,
                  force         = 100,
                  segment.size  = 0.2,
                  segment.color = "grey50",
                  direction     = "x") +
  geom_text_repel(data          = subset(df, Rank <= 5),
                  nudge_y       = 32 - subset(df, Rank <= 5)$Rank,
                  size          = 4,
                  box.padding   = 1.5,
                  point.padding = 0.5,
                  force         = 100,
                  segment.size  = 0.2,
                  segment.color = "grey50",
                  direction     = "x")
p <- p + labs(title =  "Country Tax Haven Rankings", x = "Country Rank", y = "Tax Haven Score")
p <- p + theme(plot.title  = element_text(hjust = 0.5, vjust = .1),
               text=element_text(size=16,  family="serif"), 
               panel.grid.major.y = element_line(color = 'salmon'),
               panel.grid.major.x = element_blank(),
               axis.ticks.x.bottom = element_blank(),
               axis.ticks.y = element_blank(),
               axis.text.x = element_blank(),
               panel.background = element_rect(fill='transparent'))
plot(p)

```

---
class: center, middle
### Secrecy Index
```{r, echo=FALSE, warnings = FALSE, out.width= "60%", out.height= "40%"}
g <- ggplot(data = df3, aes(x = `FSI Rank`, y = `FSI Score`, label = ISO.2)) +
  geom_point(color = case_when(
    df3$`FSI Rank` >= 135 ~ 'dodgerblue',
    df3$`FSI Rank` <= 5 ~ 'forestgreen',
    TRUE ~ 'slateblue1'), 
    size = 1, alpha = 0.8) +
  geom_text_repel(data          = subset(df3, `FSI Rank` >= 137),
                  nudge_y =     500 - subset(df3, `FSI Rank` >= 137)$`FSI Rank`,
                  size          = 4,
                  box.padding   = 1.5,
                  point.padding = 0.5,
                  force         = 50,
                  segment.size  = 0.2,
                  segment.color = "dodgerblue",
                  direction     = "x") +
  geom_text_repel(data          = subset(df3, `FSI Rank` <= 5),
                  nudge_y       = 500 - subset(df3, `FSI Rank` <=5)$`FSI Rank`,
                  size          = 4,
                  box.padding   = 1.5,
                  point.padding = 0.5,
                  force         = 100,
                  segment.size  = 0.2,
                  segment.color = "forestgreen",
                  direction     = "x") +
  scale_x_continuous(
    limits = c(0, 150, 50),
    breaks = c(30, 60 , 90, 120, 150))
options(ggrepel.max.overlaps = 15)  
    
g <- g + labs(title =  "Country Secrecy Ranking", y = "Secrecy Score", x = "Country")
g <- g + theme(plot.title  = element_text(hjust = 0.5, vjust = .1),
            axis.title.x = element_blank(),
            panel.grid.major.x = element_line(color = 'mistyrose'),
            panel.background = element_rect(fill='transparent'),
            text=element_text(size=16,  family="serif"))
plot(g)
```
* Top 5 countries with highest secrecy rating are: United States, Switzerland, Singapore, Hong Kong and Luxembourg.  
* Bottom 5 countries are Montserrat, Slovenia, Gambia, Argentina and San Marino